---
title: "Story Telling with Data (Data Visualization)"
author: "Mohammad Ali Momen"
date: "04/16/2023"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: true
    self_contained: true
    code_download: true
    code_folding: show
    df_print: paged
  md_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    variant: markdown_github
  html_notebook: default
  pdf_document: default
  word_document: default
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, attr.source = '.numberLines')
```

***

**Data Analysis methodology**: CRISP-DM

**Dataset**: Sales data of a retail company in USA

**Case Goal**: Comparing between Men & Women in each different cities, is there any difference of generating total revenue to company from gender aspect?

***

# Required Library
```{r}
library('ggplot2')
```

***


# Read Data from File
```{r}
data <- read.csv("CS_02_01.csv", header = T)
dim(data)  # 14059 records, 16 variables
head(data)
summary(data[, c("Gender", "Revenue")])
```

***

# Data PreProcessing
## Data Cleaning
```{r}
data$Revenue <- as.numeric(gsub("[$]", "", data$Revenue))
data$Gender <- factor(data$Gender)
summary(data[, c("Gender", "Revenue")])
```

## Data Preparation
```{r}
city_gender_rev_1 <- as.data.frame(tapply(data$Revenue,
                                          list(data$City, data$Gender),
                                          sum))  # total revenue in each city per each gender
city_gender_rev_1

city_gender_rev_1$total_rev <- city_gender_rev_1$M + city_gender_rev_1$F  # total revenue
city_gender_rev_1 <- city_gender_rev_1[order(city_gender_rev_1$total_rev), ] #sort dataframe based-on total_rev column in each city
city_gender_rev_1$city <- rownames(city_gender_rev_1)  # add city names
city_gender_rev_1$city <- factor(city_gender_rev_1$city, levels = reorder(city_gender_rev_1$city, city_gender_rev_1$total_rev))  # sort city names levels based-on total_rev
rownames(city_gender_rev_1) <- NULL  # remove rownames of dataframe

c1 <- city_gender_rev_1[, c("city", "F")]
colnames(c1)[2] <- "revenue"
c1$gender <- "F"
c2 <- city_gender_rev_1[, c("city", "M")]
colnames(c2)[2] <- "revenue"
c2$gender <- "M"
city_gender_rev_2 <- rbind(c1, c2)
city_gender_rev_2  # appropriate dataframe for ggplot2
```


# Plot without a story to tell
```{r}
ggplot(city_gender_rev_2, aes(y = city, x = revenue, fill = gender)) +
  geom_col(position = 'dodge')
```

# Plot with a story to tell
```{r}
ggplot(city_gender_rev_2, aes(x = revenue, y = city)) +
  geom_point(aes(color = gender)) +
  geom_line(aes(group = city))
```

Prepare dataframe for labels and highlights
```{r}
city_gender_rev_1$diff_percent <- round((pmax(city_gender_rev_1$M, city_gender_rev_1$F) / pmin(city_gender_rev_1$M, city_gender_rev_1$F) - 1) * 100, 2)
city_gender_rev_1

city_gender_rev_1$gender_max <- ifelse(city_gender_rev_1$F > city_gender_rev_1$M, "F", "M")  # maximum revenue in each city by gender
```

Identify cities with 20% or higher gap in revenue between Men and Women (cities with diff_percent >= 20)
```{r}
big_diff_cities <- city_gender_rev_1$city[city_gender_rev_1$diff_percent >= 20]
```

Create another dataframe for creating highlights and labels
```{r}
highlight <- city_gender_rev_2[city_gender_rev_2$city %in% big_diff_cities, ]
highlight$diff_percent <- city_gender_rev_1[match(highlight$city, city_gender_rev_1$city), "diff_percent"]
highlight$gender_max   <- city_gender_rev_1[match(highlight$city, city_gender_rev_1$city), "gender_max"]
highlight

plot_labels <- highlight[highlight$gender == highlight$gender_max,]  # remove duplicate rows

plot <- ggplot(city_gender_rev_2, aes(x = revenue, y = city)) +
  geom_line(aes(group = city), alpha = 0.3) +
  geom_point(aes(color = gender), size = 1.5, alpha = 0.3) +
  geom_line(data = highlight, aes(group = city)) +
  geom_point(data = highlight, aes(color = gender), size = 2) +
  geom_text(data = plot_labels, aes(color = gender_max, label = paste0("+", diff_percent, "%")), size = 3, hjust = - 0.5)

plot
```

Final plot
```{r}
plot <- plot + scale_color_discrete(labels = c("Female", "Male")) +
  scale_x_continuous(labels = scales::dollar, expand = c(0.02, 0),
                     limits = c(0, 10500),
                     breaks = seq(0, 10000, by = 2500)) +
  scale_y_discrete(expand = c(.02, 0)) +
  labs(title = "Total Revenue by City and Gender",
       subtitle = "Out of 23 cities, eight locations experience a 20% or greater difference \nin revenue generated by males versus females. Hidalgo experiences the \ngreatest difference with females generating 86% more revenue than males.") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.justification = c(0, 1), 
        legend.position = c(.1, 1.075),
        legend.background = element_blank(),
        legend.direction = "horizontal",
        plot.title = element_text(size = 20, margin = margin(b = 10)),
        plot.subtitle = element_text(size = 10, color = "darkslategrey", margin = margin(b = 25)),
        plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey70", hjust = 0))
plot
```

***

For more information check the [Github](https://github.com/mamomen1996/R_CS_02) repository.